---
title: "Data Scraping"
author: "Karim"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=False, message=False}
library(tidyverse)
```

### Opening Data

```{r}
defense_2023 = read.csv("Def2023.csv")
defense_2022 = read.csv("Def2022.csv")
defense_2021 = read.csv("Def2021.csv")
defense_2020 = read.csv("Def2020.csv")
defense_2019 = read.csv("Def2019.csv")
defense_2018 = read.csv("Def2018.csv")

attack_2023 = read.csv("Att2023.csv")
attack_2022 = read.csv("Att2022.csv")
attack_2021 = read.csv("Att2021.csv")
attack_2020 = read.csv("Att2020.csv")
attack_2019 = read.csv("Att2019.csv")
attack_2018 = read.csv("Att2018.csv")

pass_2023 = read.csv("Pas2023.csv")
pass_2022 = read.csv("Pas2022.csv")
pass_2021 = read.csv("Pas2021.csv")
pass_2020 = read.csv("Pas2020.csv")
pass_2019 = read.csv("Pas2019.csv")
pass_2018 = read.csv("Pas2018.csv")
```

```{r}
#Data Row for Missing Data
missing = data.frame(a = c("Tottenham Hotspur", "Tottenham Hotspur"),
                b = c(3, 3),
                c = c("Crystal Palace", "Crystal Palace"),
                d = c(0, 0),
                f = c(2, 5), 
                g = c(15/23, 10/18),
                i = c(16, 12))

colnames(missing) = c("h_team", "h_score", "a_team", "a_score", "shots_blocked", "tackle_success", "pass_blocked")

```

### Function to clean defense data

```{r}
clean_data = function(def){
  df_new = def %>%
    select(c(5, 7, 12, 14, 21, 23, 24, 33, 34, 35)) %>%
    mutate(tackle_success = TklW_Tackles / Tkl_Tackles,
           pass_blocked = Pass_Blocks + Int
    ) %>%
    select(-c(5:7, 9, 10)) %>%
    rename(c(h_team = Home_Team, 
            a_team = Away_Team,
            h_score = Home_Score, 
            a_score = Away_Score,
            shots_blocked = Sh_Blocks))
  return(df_new)
}
```

### Function to flip attack and pass rows

```{r}
swap_index = function(v){
  n = length(v)
  odd_indices = seq(1, n, by = 2)
  even_indices = seq(2, n, by = 2)
  v_swapped = v
  v_swapped[odd_indices] = v[even_indices]
  v_swapped[even_indices] = v[odd_indices]
  return(v_swapped)
}
```

```{r}
outcome = function(score1, score2){
  outcome = c()
  for (i in 1:length(score1)){
    if (score1[i] > score2[i]){
      outcome = c(outcome, "W")
      } else if (score1[i] == score2[i]){
        outcome = c(outcome, "D")
        } else {
          outcome = c(outcome, "L")
          }
  }
  return(outcome)
}
```



### Function to combine rows with "home" and "away" data for the same game.

```{r}
merge_rows = function(df){
  df_odd = df[seq(1, nrow(df), by = 2), ]
  df_even = df[seq(2, nrow(df), by = 2), ]
  
  df_odd = df_odd %>%
    mutate(outcome = outcome(h_score, a_score)) %>%
    rename(team = h_team) %>%
    select(c(1, 5, 6, 7, 8))
  
  df_even = df_even %>%
    mutate(outcome = outcome(a_score, h_score)) %>%
    rename(team = a_team) %>%
    select(c(3, 5, 6, 7, 8))

  df_new = data.frame(matrix(ncol = ncol(df_odd), nrow = nrow(df)))
  df_new[seq(1, nrow(df), by = 2), ] = df_odd
  df_new[seq(2, nrow(df), by = 2), ] =  df_even
  
  colnames(df_new) = c("team", "shots_blocked", "tackle_success", "passes_blocked", "outcome") 
  
  return(df_new)
}
```


### Creating a list of defensive, attacking, and passing data for 6 years.

```{r}
def = list(defense_2023, defense_2022, defense_2021, defense_2020, defense_2019, defense_2018)
att = list(attack_2023, attack_2022, attack_2021, attack_2020, attack_2019, attack_2018)
pas = list(pass_2023, pass_2022, pass_2021, pass_2020, pass_2019, pass_2018)
```

### Main()

#### Cleaning defensive data to have only data I need

```{r}
clean_d = list()
for (i in 1:6){
  df = clean_data(def[[i]])
  clean_d = append(clean_d, list(df))
}
```

#### Inserting the missing data for defense_2022

```{r}
clean_d[[2]] = rbind(clean_d[[2]][1:348,], missing, clean_d[[2]][-c(1:348),])
rownames(clean_d[[2]]) = NULL
```

#### Using Attack and Pass data to calculate the ratios of shots and passes blocked.

```{r}
clean = list()
for (i in 1:6){
  df = clean_d[[i]] %>%
    mutate(shots_blocked = shots_blocked / swap_index(att[[i]]$Sh),
           pass_blocked = pass_blocked / swap_index(pas[[i]]$Att_Total)
    )
  df_rows_combined = merge_rows(df)
  df_rows_combined$outcome <- factor(df_rows_combined$outcome)
  clean = append(clean, list(df_rows_combined))
}

View(clean[[1]])
```

#### Then, merging rows of each data set.

#### Merging all rows

```{r}
final = bind_rows(clean[[1]], clean[[2]], clean[[3]], clean[[4]], clean[[5]], clean[[6]]) 

rows_with_na <- final[is.na(final$shots_blocked), ]

final$shots_blocked <- ifelse(is.na(final$shots_blocked), 0, final$shots_blocked)
```

#### Saving data as csv

```{r}
write.csv(final, "C:\\Users\\kerim\\OneDrive\\Desktop\\GEO 322E\\Paper Analysis\\final2.csv", row.names=TRUE)
```
